---
title: "lab_12"
author: "derek willis"
date: "11/16/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## You will need

* A Census API key
* Our usual libraries, plus the geographic ones

## Load libraries and establish settings

**Task** Create a codeblock and load appropriate packages and settings for this lab.

```{r}
# Turn off scientific notation
options(scipen=999)
library(tidycensus)
library(tidyverse)
library(janitor)
library(lubridate)
library(ggrepel)
library(sf)
library(stringr)
library(tigris)
```

## Questions

**Q1.** You are interested in tracking police incidents around the West Virginia University campus over time, with the goal of surfacing interesting incidents and patterns. In particular, you want to know whether the latest incident titles are relatively new/novel compared to previous reports and whether an incident has happened before in the past 30 days.

Using [CSV data](https://raw.githubusercontent.com/dwillis/wvu-projects/master/crime-log/crime_log.csv), making the column names more readable and ensuring each datetime is separated into a date column and an hour column and also create a day_of_week column and populate that using lubridate.

Then write code that will construct a paragraph about the data described above comparing the most recent data to the previous date and the most recent previous date for the same day of the week. To do that, you will need to produce:

1. A dataframe with the crime log data loaded and the datetime column converted to the proper datatype.
2. Columns in your dataframe that contain the date, time and day of the week that each date represents (you may need to google how to do that).
3. Generate a dataframe of the most recent date's incidents.
4. Using the first title in that dataframe, find the number of incidents that occurred in the previous 30 days and write out a sentence using paste. The sentence should include the most recent date, the title of the incident and the number of times that kind of incident has occurred in the previous 30 days.

Put your sentence in the answer below, along with one other question you could ask of this data.

**A1.** The sentence that I generated is "There have been 3 instances of ASSIST POLICE POLICE  in the last thirty days. The last incident was 2022-04-21". Another question I could ask of this data is what crime response is the most prevalent on campus and what time of day this crime response takes place most often. 

```{r}

crime_logs_wv <- read_csv("https://raw.githubusercontent.com/dwillis/wvu-projects/master/crime-log/crime_log.csv") %>%
  clean_names() %>%
  mutate(new_date = mdy_hm(datetime)) %>%
  mutate(date = date(new_date)) %>%
  mutate(hour = hour(new_date)) %>%
  mutate(weekday = wday(new_date)) %>%
  arrange(desc(new_date))

most_recent_date <- crime_logs_wv %>%
  filter(date == "2022-04-21")

last_thirty_days <- crime_logs_wv %>%
  filter(date > "2022-03-26") %>%
  filter(title == "ASSIST POLICE POLICE")

summary_thirty_days <- last_thirty_days %>%
  group_by(title) %>%
  summarise(
    count = n()) 

sub <- paste("There have been", format(summary_thirty_days$count), "instances of", format(summary_thirty_days$title), " in the last thirty days. The last incident was", last_thirty_days$date[[1]])

print(sub)
```

**Q2.** Maryland produces a [CSV file](https://raw.githubusercontent.com/state-of-maryland/VaccineCSVs/master/Vax_ZIP_CODE_FirstSecondSingle_Totals_Raw.csv) showing the total number of vaccinations by zip code and -- interestingly -- it includes non-Maryland zip codes (as well as some zip codes that seem to be bad data). 

Write a script that reads the CSV file, making the columns more readable and adding today's date. 

Then output the top 25 Maryland zip codes and top 25 non-Maryland zip codes based on number of people fully vaccinated, and write both dataframes out to CSV files, using the following name formats:

Maryland zips: maryland_zips_yyyymmdd.csv
Non-Maryland zips: non_maryland_zips_yyyymmdd.csv

Where yymmdd is from today's date. You MUST do this programmatically so that it can run every day without being updated. Be sure to add those two files to your GitHub repository.

To do that, you will need to:

1. Get rid of rows that do not have data for fully vaccinated people *DONE*
2. Remove any rows where the zip code is less than five digits (there are several ways you could do this). *DONE*
3. Figure out the "lowest" and "highest" Maryland zip codes. *DONE*

Google the top 2-3 zip codes from both Maryland and non-Maryland results and describe them in your answer.

**A2.** The top three ZIP codes in Maryland are located in Silver Spring, Gaithersburg and Germantown. These three locations are all suburban hubs in Montgomery County and very close to one another. It makes me wonder if Montgomery County had a better vaccination rollout than other parts of the state. These three places are also densely populated compared to other areas in the state, which could also account for why these areas have so many vaccinated people. The top three ZIP codes outside of Maryland are all located in Washington, D.C. This makes me wonder about how D.C.'s vaccine rollout succeeded so much that its relatively small, yet dense population outnumbers the number of vaccinated people in even bigger cities. 

```{r}

zip_code_vax <- read_csv("https://raw.githubusercontent.com/state-of-maryland/VaccineCSVs/master/Vax_ZIP_CODE_FirstSecondSingle_Totals_Raw.csv")

clean_zip_code_vax <- zip_code_vax %>%
  filter(!is.na(FullyVaccinated)) %>%
  clean_names() %>%
  mutate(
    today_date = today(tzone = "")
  )

final_clean_zips <- clean_zip_code_vax %>%
  mutate(zip = str_sub(zip_code, start=1L, end=5L)) %>%
  filter(nchar(zip) == 5)

zip_codes_maryland <- final_clean_zips %>%
  filter(zip >= 20588, zip <= 21930) %>%
  group_by(fully_vaccinated) %>%
  arrange(desc(fully_vaccinated)) %>%
  head(25)

zip_codes_other_states <- final_clean_zips %>%
  filter(zip < 20588 | zip > 21930) %>%
  group_by(fully_vaccinated) %>%
  arrange(desc(fully_vaccinated)) %>%
  head(25)

write_csv(zip_codes_maryland, file = paste0('maryland_zips_', Sys.Date(), '.csv'))

write_csv(zip_codes_other_states, file = paste0('non_maryland_zips_', Sys.Date(), '.csv'))

```

**Q3.** Maryland has had one of the more successful vaccination rates, but you want to understand where a significant portion of the population remains less than fully vaccinated. 

Make a dataframe of all Maryland zip codes from the vaccination data you used in Q2 and join it to Maryland zip code population data from the Census API. *DONE*

Calculate the percentage of fully vaccinated for each zip code. *DONE*

Then make a map showing the zip codes with less than 50 percent fully vaccinated. *DONE*

Put a blank map with the outline of Maryland counties (which you can get from Tigris) behind it for context. *DONE* 

Add a label to the map describing what the data is and include the date from the vaccination data (programmatically, so you can run it every day). *DONE*

You WILL need to ensure that the columns you use to join the two dataframes are the same datatype and remember that join order matters. *DONE*

Describe the output of the map below, and say what you think the most interesting areas are.

**A3.** The map outputs ZIP codes that are mostly concentrated in the western part of the state near the border with Pennsylvania and West Virginia. A majority of these ZIP codes have between 30-50% of their populations vaccinated. This makes me wonder if the vaccination rollout in this part of the state is lacking, or if the political orientation of these ZIP codes and counties had a lot to do with people's choices to get vaccinated or not. Nearly every ZIP code that shows up on this map is in a part of the state that is more rural. Something that stands out to me about this map is how there is a chunk of unvaccinated individuals smack dab in the middle of Prince George's County, which is strange because the county overall has a high vaccination rate. The populations where between 40-50 percent of the population are vaccinated are concentrated in rural Maryland, both the western edge and the eastern shore. There are large clusters in Caroline, Dorchester and Cecil counties, all of which are rural areas in the state. 

```{r}

census_api_key("102c0c72d8f8eb880608f9dacb53fdf9100f8ba0")

acs2019 <- load_variables(2019, "acs5", cache = TRUE)
View(acs2019)

md_pop <- get_acs(
  geography = "zcta", 
  variables = c(name = "B02001_001"),
  year = 2019,
  geometry = TRUE
)

three_zip_codes_maryland <- final_clean_zips %>%
  filter(zip >= 20588, zip <= 21930) %>%
  group_by(fully_vaccinated) %>%
  arrange(desc(fully_vaccinated))

combined_vax_zip <- md_pop %>%
  inner_join(three_zip_codes_maryland, by=c("GEOID" = "zip")) %>%
  rename(population=estimate) %>%
  select(GEOID, population, fully_vaccinated, geometry)

map_vax <- combined_vax_zip %>%
  mutate(percent = (fully_vaccinated/population)*100) %>%
  filter(percent < 50) %>%
  arrange(desc(percent)) %>%
  select(GEOID, percent)

counties <- counties() 

md_counties <- counties %>%
  filter(STATEFP == "24")
  
md_counties %>%
  ggplot() + 
  geom_sf() + 
  theme_minimal()

ggplot() +
  geom_sf(data=md_counties) +
  geom_sf(data=map_vax, aes(fill=percent)) +
  scale_fill_viridis_c(option = "plasma", trans = "sqrt") +
  theme_void() +
  labs(title = paste("Maryland ZIP codes with less than 50 percent of population vaccinated as of ", current[[1]], sep=""))

```
