---
title: "Class Reference"
author: "Student name"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

Consider this a personal guide to the commands and functions you will learn. In general, when you come across an R command or function that you want to remember, put it in here along with a description of what it does and when you'd use it.

## Things I learned on Jan. 27

We won't do anything in this class that doesn't contain the command library(tidyverse)
The command to assign a name to a variable is number <- 2
The command to set a working directory is setwd(). For example:

## How to find columns where the value is NA

filter(is.na(forgiveness_date)) %>% 

## Columns that are not NA

  filter(!is.na(forgiveness_date)) %>%
  
## Changing column names in the data set

cleaned_arnold_md_loans <- arnold_md_loans %>%
  clean_names() %>%
  rename(id = x1_id) 


```{r}
setwd("~/Documents/GitHub/data_journalism_2022_spring")
```

```{r}
install.packages("tidyverse")
library(tidyverse)
```