---
title: "Class Reference"
author: "Student name"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

Consider this a personal guide to the commands and functions you will learn. In general, when you come across an R command or function that you want to remember, put it in here along with a description of what it does and when you'd use it.

## Things I learned on Jan. 27

We won't do anything in this class that doesn't contain the command library(tidyverse)
The command to assign a name to a variable is number <- 2
The command to set a working directory is setwd(). For example:

## How to find columns where the value is NA

filter(is.na(forgiveness_date)) %>% 

## Columns that are not NA

  filter(!is.na(forgiveness_date)) %>%
  
## Changing column names in the data set

cleaned_arnold_md_loans <- arnold_md_loans %>%
  clean_names() %>%
  rename(id = x1_id) 
  
## More notes on Janitor to prepare for lab 3

If you're going to describe all of the data based on ZIP code, for example, it needs to be on the same basis
  5 digit basis because all of the records have at least five digits
  Reduce to least common denominator ish
  
  glimpse is very helpful to know what character types you're working with
    if it says a number is a 
    
    clean_names()
      "whatever data set I'm working with, just save the names and make them better
      
      rename (old, wrong name -> new name)
      
      distinct()
        gets rid of duplicates
        retrieves every unique row that is present
        
        
  mutate(zip = str_sub(zip, start=1L, end=5L))
    str_sub takes a subset of the character column 
      takes what's in the zip column, start with the first character from the left and end with the fifth character from the left
      
      all changes to columns have to be in a separate mutate function
      
      str_to_title
        standardizes the city names to be title case
        doesn't matter if it's uppercase, etc it just has to be consistent
```{r} 
      # making consistent spellings for cities
        mutate(city = case_when (
        city == "Anold" ~ "Arnold",
        TRUE ~ city
        ))

``` 

```{r}
setwd("~/Documents/GitHub/data_journalism_2022_spring")
```

```{r}
install.packages("tidyverse")
library(tidyverse)
```