---
title: "Class Reference"
author: "Student name"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

install.packages("tidycensus")
```

## Introduction

Consider this a personal guide to the commands and functions you will learn. In general, when you come across an R command or function that you want to remember, put it in here along with a description of what it does and when you'd use it.

## Things I learned on Jan. 27

We won't do anything in this class that doesn't contain the command library(tidyverse)
The command to assign a name to a variable is number <- 2
The command to set a working directory is setwd(). For example:

## How to find columns where the value is NA

filter(is.na(forgiveness_date)) %>% 

## Columns that are not NA

  filter(!is.na(forgiveness_date)) %>%
  
## Changing column names in the data set

cleaned_arnold_md_loans <- arnold_md_loans %>%
  clean_names() %>%
  rename(id = x1_id) 
  
## More notes on Janitor to prepare for lab 3

If you're going to describe all of the data based on ZIP code, for example, it needs to be on the same basis
  5 digit basis because all of the records have at least five digits
  Reduce to least common denominator ish
  
  glimpse is very helpful to know what character types you're working with
    if it says a number is a 
    
    clean_names()
      "whatever data set I'm working with, just save the names and make them better
      
      rename (old, wrong name -> new name)
      
      distinct()
        gets rid of duplicates
        retrieves every unique row that is present
        
        
  mutate(zip = str_sub(zip, start=1L, end=5L))
    str_sub takes a subset of the character column 
      takes what's in the zip column, start with the first character from the left and end with the fifth character from the left
      
      all changes to columns have to be in a separate mutate function
      
      str_to_title
        standardizes the city names to be title case
        doesn't matter if it's uppercase, etc it just has to be consistent
```{r} 
      # making consistent spellings for cities
        mutate(city = case_when (
        city == "Anold" ~ "Arnold",
        TRUE ~ city
        ))

        mutate(city = case_when (
        str_detect(city, "^Arnl") ~ "Arnold"
            #string detect = if you find that the city name starts with "Arnl" then change it to Arnold
            #using a pattern, rather than "city ==" is a great way to filter en masse
        TRUE ~ city
        ))

``` 

## Order of operations for filter, mutate, group by, summarise

Filter -> first because it should narrow down as much data as you can before starting to mutate names, etc.
mutate -> second because this is how the data gets narrowed down based on name, misspelling etc.
    within mutate, you should do a case_when first and then do str_detect first
    each line is getting executed in the order that it appears, so
group_by, summarise, then arrange 

Example:

## How to manipulate a data frame - Solid Template

```{r}

my_result <- original_data %>%
  filter(some_column == "some_value") %>%
  clean_names() %>%
  mutate(some_column = str_to_title(some_column)) %>%
  group_by(some_column)  %>%
  summarize(new_aggregate = n()) %>% OR summarise(new_aggregate = sum(some_number_column)) %>%
  arrange(desc(new_aggregate)) 


{r}
```

## Filter Tags

& = data must be this AND this
| = data must be this OR this

filter = when there is a specific criteria 
group_by and summarize = we want totals
mutate = makes the data fit a specific pattern

## Bind Rows

All three of these datasets have the same number of columns, all with the same names, so if we want to merge them together to compare them over time, we need to stack them together. The verb here, in R, is bind_rows
    bind_rows expects every single column, including cases, to be identical
    it also expects every single column to be a different data type
    
## Joins

Start with original dataframe %>% left_join(naics_codes, by="naics_code")

## To make a csv file from other data that you have imported

write_csv(name of data frame I'm writing, "data/full path name of cvs file (name this the same thing as dataframe.csv") 
AKA
write_csv(college_park_loans, "data/college_park_loans.csv")

## Changing the column names
lender_types <- read_csv("tabula-PPP_Report_Public_210531-508.csv", skip=1, col_names=c("type", "count", "approved", "net_dollars"))

"skip=1 is skipping the first row, after this rename the column names using col_names=c"

```{r}
install.packages("tidyverse")
library(tidyverse)
setwd("~/Documents/GitHub/data_journalism_2022_spring")


all_payouts_decade %>%
  filter(department == "Police Department" & fiscal_year == 2017) %>%
  summarise(total = sum(amount))


```