---
title: "lab_03"
author: "sean mussenden"
date: "8/19/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries and establish settings
```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse here

library(tidyverse)

# Load janitor here

library(janitor)

```

## Load data
```{r}
# Write code to load the WV slice of the ppp loan data here and call it wv_ppp_loans.
# The data is called lab_03.rds, and is in the same folder as lab_03.Rmd (this data notebook)

wv_ppp_loans <- read_rds("lab_03.rds")

```

## Answer questions

Q1: Which five cities located in Kanawha County, West Virginia, got the largest _number_ of loans? How many did each get? To answer this question, you'll have to clean the very messy "city" field.

Note: To answer this question, you'll need to see every variation of the "city" column in Kanawha County, and then fully clean the city field for those cities. Hint: once cleaned, the dataframe that counts the number of loans by city should have 38 distinct cities for Kanawha County.

A1: The cities that got the largest number of loans in Kanawha County are Charleston, South Charleston, Saint Albans, Dunbar and Nitro.

```{r}

# Write your code in this codeblock.

cleaned_cities <- wv_ppp_loans %>%
  filter(project_county_name == "KANAWHA") %>%
  clean_names() %>%
  mutate(city = str_to_title(city)) %>%
  mutate(city = case_when(
    city == "S Charleston" ~ "South Charleston",
    city == "So Charleston" ~ "South Charleston",
    city == "Charelston" ~ "Charleston",
    city == "Balle" ~ "Belle",
    city == "Crosslanes" ~ "Cross Lanes",
    city == "Elkveiw" ~ "Elkview",
    str_detect(city,"^St.") ~ "Saint Albans",
    TRUE ~ city
  )) %>%
  group_by(city)  %>%
  summarize(
    count_loans = n()) %>%
  arrange(desc(count_loans)) 

```

Q2: Find the 5-digit ZIP code in West Virginia that had the highest total amount of loans (by dollar value). What is the name of the business that got the largest loan (by dollar value) in that zip code? Write up a short paragraph that also answers these questions:

* Where in West Virginia is this ZIP code located?
* How much was the business' loan for?
* How much of that loan was forgiven?
* Which number is bigger, the loan amount or the forgiveness amount? Can you explain the discrepancy?

A2: The ZIP code is 25301 and it is located in Charleston, West Virginia, which is the capital of the state. The business with the biggest loan in this ZIP code is The Eye and Ear Clinic of Charleston. Their loan amount was $1520007. 

```{r}
# Hint
# First, create a dataframe that shows the top 5-digit ZIP code by total dollar amount

cleaned_zip <- wv_ppp_loans %>%
    mutate(zip = str_sub(zip, start=1L, end=5L)) %>%
    group_by(zip) %>%
      summarise(total_amount = sum(amount)) %>% 
      arrange(desc(total_amount))


# Second, use the top ZIP code to create a separate dataframe that shows the business with the biggest loan in that ZIP code.

zip_businesses <- wv_ppp_loans %>% 
  filter(zip == "25301") %>%
  arrange(desc(amount))
  

```

Q3: Five business entities owned by the spouse of a member of West Virginia's U.S. House delegation got PPP loans. Write up a paragraph that answers the following questions.   

  * What are the businesses' names?
  * What do the businesses do?
  * How much money did the five businesses receive in total?
  * What percentage of the total loan amount have been forgiven?
  * How did the U.S. House member vote on the legislation initially authorizing the PPP loan program.

A3:The business' are owned by the spouse of Carol Miller. The business names are Dutch Miller Chevrolet, Dutch Miller of Charleston, Dutch Miller of Charlotte, Dutch Miller Subaru, and DM Motors. All of the businesses are car dealerships. The businesses received $3,105,300 in total loans. The amount of the total loans that were forgiven was 3,140,471. With that said, the percent forgiven is 101.1326. Miller initially voted in favor of the legislation that authorized the program. She was also a co-sponsor of the bill. 

```{r}

# Hint
# First, create a dataframe that only has loans to those five companies

miller_loans <- wv_ppp_loans %>%
  filter(str_detect(name, "DUTCH MILLER |DM MOTOR"))
  

# Second, use that newly created dataframe to calculate total loan amount and percentage forgiven for those companies

miller_loans %>%
  mutate(total_loan_amount = sum(amount),
        total_forgiveness = sum(forgiveness_amount),
        percent_forgiveness = (total_forgiveness/total_loan_amount)*100) %>%
  select(name, amount, forgiveness_amount, total_loan_amount, total_forgiveness, percent_forgiveness)
  
  
```