---
title: "lab_11"
author: "sean mussenden"
date: "8/26/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## You will need

* A Census API key

## Load libraries and establish settings

**Task** Create a codeblock and load appropriate packages and settings for this lab. You'll need tidyverse, tidycensus, janitor, sf and tigris, and you'll also need to install your census API key.

```{r}
library(tidyverse)
library(sf)
library(janitor)
library(tigris)
library(lubridate)
library(tidycensus)
census_api_key("549950d36c22ff16455fe196bbbd01d63cfbe6cf")
```

## Load data

**Task** Create a codeblock and load the WV slice of the PPP loan data, which is in the data folder and is called lab_11.rds.

```{r}
wv_ppp_loans <- read_rds("data/lab_11.rds")

```

## Questions

**Q1.** You are interested in doing some reporting on how West Virginia companies in the "Mining, Quarrying, and Oil and Gas Extraction" industry (https://www.naics.com/six-digit-naics/?code=21) used PPP loans during the pandemic (West Virginia is a big coal mining state).  

Then produce a choropleth map that shows West Virginia counties shaded according to the total amount of loans to that industry per 100,000 population.

Use the magma color palette, and shade on a "log" scale. Also produce a table that shows the total amount of loans per 100,000 population for each county, ranked from highest to lowest, along with the total amount of loans and the total number of loans.

```{r}
wv_mining <- wv_ppp_loans %>%
  mutate(naics_clean = str_sub(naics_code, start=1L, end=2L)) %>%
  filter(naics_clean == 21)

wv_county_population <- get_acs(geography = "county", 
  variables = c(population = "B01001_001"), 
  state = "WV",
  geometry = TRUE)

wv_mining

wv_county_population

wv_county_population_clean <- wv_county_population %>%
  rename(COUNTY = NAME) %>%
  mutate(COUNTY = toupper(str_remove_all(COUNTY,", West Virginia|County"))) %>%
  mutate(COUNTY = str_trim(COUNTY,side="both")) %>%
  rename(population = estimate) %>%
  select(COUNTY, population, geometry)

wv_mining_clean <- wv_mining %>%
  group_by(project_county_name) %>%
  summarise(
    count=n(),
    total = sum(amount)
  ) 

loans_per_100k <- wv_county_population_clean %>%
  left_join(wv_mining_clean, by=c("COUNTY" = "project_county_name")) %>%
  mutate(loans_per_100k = total/population*100000) %>%
  mutate(loans_per_100k = replace_na(loans_per_100k,0)) %>%
  arrange(desc(loans_per_100k)) 

  ggplot() + 
  geom_sf(data=loans_per_100k, aes(fill=loans_per_100k)) + 
  theme_minimal() +
  scale_fill_viridis_b(option="magma",trans = "log") 

```

Then answer these questions:

* Based on the map, what region of the state appears to have gotten more loans in this industry? What geographic pattern do you notice in the data?
* Based on the table, what county got the highest total loan amount (based on 100k population)? What was the total amount per 100K population, the total amount of loans (not population adjusted), how many loans were there, and what's the population?

**A1.** The region of the state that appears to have the highest total loan amount is the southwestern and northwestern parts of the state. There is a geographic pattern in the data that suggests the mining industry is a main source of the western part of the state's economy, while the grey parts of the state that do not have any data are concentrated in the eastern part of the state. The county with the highest total loan amount per 100k people was Ritchie county. The total loan amount per the 100k population was 77752742.79. The total amount of loans was 7653980.00, there were 9 loans, and the population is 9844.

**Q2.** You want to understand WHY the spatial pattern you saw above exists.  So you grab a shapefile with the latitude and longitude coordinates of every mine permit in West Virginia from the state Department of Environmental Protection.

Here's the url: https://tagis.dep.wv.gov/arcgis/rest/services/WVDEP_enterprise/mining_reclamation/MapServer/0/query?where=1%3D1&outFields=*&geometry=&geometryType=esriGeometryEnvelope&inSR=4326&spatialRel=esriSpatialRelIntersects&outSR=4326&f=json

(It's in a slightly different format than the .shp file in the lab. It's another type of shapefile called "geojson".  It will work just the same, though, despite that difference. You can read it directly into R without manually downloading it, by putting the url in the st_read() function we learned about in this week's prelab.)

This data includes both currently operating mines and old ones that have closed.  You want to examine only mines that have a permit expiration date after Sept. 1, 2021.

Make a map with a point for each one of these mine permits and put a blank map with the outline of West Virginia counties (which you can get from Tigris) behind it.

What spatial pattern do you see? How does it compare to the spatial pattern in the first map you made?

```{r}

wv_shapefile <- st_read("https://tagis.dep.wv.gov/arcgis/rest/services/WVDEP_enterprise/mining_reclamation/MapServer/0/query?where=1%3D1&outFields=*&geometry=&geometryType=esriGeometryEnvelope&inSR=4326&spatialRel=esriSpatialRelIntersects&outSR=4326&f=json")

wv_expirations_clean <- wv_shapefile %>% 
  mutate(expiredate = ymd(expiredate)) %>%
  filter(expiredate > "2021-09-1")

wv_expirations_clean %>%
  ggplot() + 
  geom_sf() + 
  theme_minimal()

counties <- counties() 

wv_counties <- counties %>%
  filter(STATEFP == "54")

wv_counties %>%
  ggplot() + 
  geom_sf() + 
  theme_minimal()

ggplot() + 
  geom_sf(data=wv_counties) + 
  geom_sf(data=wv_expirations_clean) +
  theme_minimal()
```

**A2.** Similar to the first map I made, the active mines in this map are concentrated in the southwestern part of the state. There are so many black dots on this map that some counties almost look consumed by the mines present within them. It generally looks like the counties that had high amounts of PPP loans are the counties with the highest number of mines in them, with one notable outlier in the northwestern part of the state that only has one mine yet the highest amount of PPP loans for mines allocated to it. 

**Q3.** To confirm the spatial comparison you identified in question 2, let's create a new map that layers the mine location points on top of our PPP choropleth map we created in question 1. What do you notice about the number of mines in the county that has the highest total loan amount per 100K?

```{r}

ggplot() + 
  geom_sf(data=loans_per_100k, aes(fill=loans_per_100k)) +
  theme_minimal() +
  scale_fill_viridis_b(option="magma",trans = "log") +
  geom_sf(data=wv_expirations_clean) 

```

**A3.** There is only one mine that is in the county with the highest loan amount per 100k people. The majority of the mines in the state are located where the moderate to high loan amounts were allocated, which is mainly in the southwestern part of the state. This map makes me wonder why such a high amount of PPP loans that were allocated to mining in the state were allocated to a county with no mines and how the general mining industry in the state operates. 

**Q4.** Starting from the original WV ppp dataframe, examine the loans to companies in "Mining, Quarrying, and Oil and Gas Extraction" in the county with the highest total loan amount per 100K (you can use the table you created in question 1 to identify the top county).

```{r}

ritchie_loans <- wv_ppp_loans %>%
  mutate(naics_clean = str_sub(naics_code, start=1L, end=2L)) %>%
  filter(naics_clean == 21) %>%
  filter(project_county_name == "RITCHIE") %>%
  arrange(desc(amount))

```

* What single company accounts for the biggest chunk of the total loan amount?
* What specific industry are they in, based on their 6-digit NAICS code? Look them up on the web. What services do they provide?
* How many jobs did they save, according to the data?
* What's your hypothesis for why a company in their specific industry would need to seek a PPP loan during the pandemic?

**A4.** Hall Drilling LLC is the company that accounts for the biggest chunk of the total loan amount. Their NAICS code is for the drilling oil and gas wells sub-industry and it lists the description as "support activities for mining" on Open Corporates. The loan amount is 6866600 and the data says that the loan saved 387 jobs. Online, this company's website says that it is "one of the most respected oil & gas companies in the Appalachian Basin with over 300 skilled industry employees." Hall Drilling, LLC has 15 total employees across all of its locations and generates $6.46 million in sales and describes it as a "support activity for mining."
