---
title: "lab_05"
author: "Sean Mussenden"
date: "8/24/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## About this lab

To complete this lab, you need to:
* run existing code as directed (look for **Task**).
* modify existing code as directed (look for **Task**).
* write code in empty codeblocks provided to answer questions included (look for **Q**).
* write out the answer in the form of a complete sentence in the space given (look for **A**).

When you are finished, commit changes and push to your personal GitHub repo, then submit the URL to this document on ELMS.

## Load libraries and establish settings

You'll need to load two packages for this: the tidyverse and janitor.

**Task** load these two packages.

```{r}
# Turn off scientific notation
options(scipen=999)
library(tidyverse)
library(janitor)

```

## Load Data

You'll need to load three data sets for this:

* The West Virginia slice of the PPP loan data (lab_05.rds).
* A "lookup table" that allows you to translate NAICS (industry) numeric codes to industry titles (naics_codes.csv).
* A table of West Virginia population by county (American Community Survey, 2019 5-year averages) (wv_population_county.csv).

All three data sets are in the data folder.  Write code to load the three in the codeblock below.

**Task** Create a codeblock below this task, then read the data sets in in and assign them to appropriate variable names. There's a fourth data set you'll use in this lab, too, of selected loans in ZIP Code 25401. But there's no need to load it now.

```{r}

naics_codes <- read_csv("data/naics_codes.csv")
wv_population_county <- read_csv("data/wv_population_county.csv")
wv_ppp_loans <- read_rds("data/lab_05.rds")
zip_25401_loan_sample <- read_csv("data/zip_25401_loan_sample.csv")

```

## Answer questions

**Q1.** In the data folder, there is a csv called zip_25401_loan_sample.csv.  It contains a sample of loans from West Virginia ZIP Code 25401.

As we read earlier this semester, [multiple loan applications coming from multiple businesses at the same residential street address](https://www.nytimes.com/2021/08/17/business/ppp-fraud-covid.html) might point to fraud. Or it could alert us to companies that used [multiple corporate entities to get more money than envisioned](https://www.propublica.org/article/different-names-same-address-how-big-businesses-got-government-loans-meant-for-small-businesses) by the original law that authorized the program.   

You are going to examine this data to see if we can find a lot of loans coming from the same address.  Here's the problem: the street address field is pretty messy.  The same address appears with minor variations --  "1003 Sushruta Dr" vs "1003 SUSHRUTA DR" -- that will prevent proper grouping.

First, upload the data into Open Refine and standardize/clean the address field - make sure to make a copy of that column to work on and to try all of the options for clustering. If you've done it properly, you should have 65 discrete addresses.

Then export the data from Open Refine, and move it to the proper folder.

Next, load it in the codeblock below, assigning it to an appropriate variable name.

Then answer these questions:
* What is the street address in this data that has the most loans?
* How many loans are there at that street address?
* What are the names of the businesses at that address?

**A1.** 126 E Burke St is the street address in this data that has the most loans. There are six loans at this street address. The names of the businesses at this street address are Drew Holdings LLC, Brix27 LLC, Abraham Ashton, Hubb Co-Op LLC and Ronnin Properties LLC.

```{r}

read_csv("data/zip_25401_loan_sample.csv") 
clean_25401 <- read_csv("data/zip_25401_loan_sample_clean.csv") 

clean_25401 %>%
  group_by(address) %>%
  summarise(
    count=n()
  ) %>%
  arrange(desc(count))

clean_25401 %>% 
  filter(str_detect(address_original, "126 ")) %>%
  select(name, amount, address, date_approved)

```

**Q2.** Do some web research on the businesses that answered question 1.  

Google the street address.  Google the business names and search for their corporate records on [Open Corporates](https://opencorporates.com/). Be sure to find the website of the name of the company that appears twice in the list of businesses and develop an understanding of what it does.

Based on your research, does it seem suspicious that this collection of businesses all got loans using the same address? Why or why not. Use specific facts identified in your research to support your position.

**A2.** I don't think it seems suspicious that these companies all got loans at the same address. The address in question is a community workspace where different companies can rent out and use the space for business or other events. It seems that Ronin Properties is a real estate agent's LLC and it is very possible that this agent works at the Hub rather than his own office space. Brix27 seems to be a restaurant that is inside of the Hub where workers can go. Abraham Ashton is a sole proprietorship, which makes it more realistic for the single owner to rent a space like the Hub for doing work. Open Corporates also lists Drew Holdings as a single-person operation that works in real estate, so this address also makes sense as a place where the owner would conduct their business operations. The only thing that seemed a little bit suspicious is that the Hub itself had two loans, but based on looking at the date_approved column for this business, it is clear that one loan was issued in the earlier days of the PPP program and one was issued in 2021.

**Q3.** Start by using the West Virginia slice of the PPP loan data that you loaded at the start of the lab to create a subset of PPP loans in West Virginia's second largest county (which you can find in wv_population_county). And then use that table you created to answer the following questions:

* Which city in that county had the highest number of loans?
* In that city, which industry title had more loans than any other industry title?

Requirement: you MUST export a dataframe of PPP loans from R Studio at some point in the process (not necessarily at the beginning!), load it into Open Refine, clean the city column, export it from Open Refine, and reimport into R Studio. To export data from R Studio, you will use the write_csv() function, which works like this if we're using the wv_population_county dataframe:

Guidance: there are a lot of steps you'll need to take to answer this question. You may find it helpful to write out in English what you plan to do step-by-step before you start writing code.   

**A3.** The county is Berkeley County. The city in the county with the highest number of loans is Martinsburg. The industry that had more loans than any other industry title is full-service restaurants. 

```{r}

wv_population_county %>%
  arrange(desc(population_2019))
  
berkeley_loans <- wv_ppp_loans %>%
  filter(project_county_name == "BERKELEY")

write_csv(berkeley_loans, "data/berkeley_loans")

berkeley_clean_cities <- read_csv("data/berkeley-loans.csv")
berkeley_clean_cities %>%
  group_by(city_edited) %>%
  summarize(
    count=n()
  ) %>%
  arrange(desc(count))

martinsburg_ppp <- berkeley_clean_cities %>%
  filter(city_edited == "Martinsburg")

martinsburg_ppp_with_naics <- martinsburg_ppp %>% left_join(naics_codes, by="naics_code")

martinsburg_ppp_with_naics %>%
  group_by(naics_code) %>%
  summarise(
    count=n()
  ) %>%
  arrange(desc(count))

```

**Q4.** What are your two best hypotheses, which you'd need to confirm with reporting and further analysis, that explain why that industry is at the top of the list?

**A4.** This industry is at the top of the list because, overall, restaurants have been some of the most affected businesses during the pandemic. Since people can no longer dine in and it is a safety hazard for emplotees to gather in person, restaurants were probably the most prominent business category to close during the pandemic. My second hypothesis is that Martinsburg, West Virginia, is a city where a lot of people/travelers/tourists pass through and therefore has a higher number of restaurants than other cities across the state. 

**Q5.** Start with a dataframe of loans to all businesses in the city and industry that answered question 3. Answer the following questions:
* What is the name of the business that got the highest approved loan amount?
* How much was it for?
* When was it approved?
* How many jobs does the data say were retained?
* Is there a difference between the business' name in the PPP data and the name its customers know it by? If so, what is that name?
* How many locations does this business have?
* Did one of its locations close during the pandemic, either before or after it got the loan?

Hint: you will not find the answers to the last three questions in the data.  You could call them directly to get that information, but I don't want you to do that for this assignment.  Instead, do some web research. I would start by Googling the company name from the data and looking at the page that comes up for the business from at http://apps.sos.wv.gov/. I would use information I found on that page and use info about the company from Google, the [Wayback machine](https://archive.org/web/) (which lets you look at older versions of a company's website), Yelp, and Facebook.

**A5.** The name of the full-service restaurant business that got the highest approved loan amount in Martinsburg, West Virginia, is COTTLE CHRISTI L LLC. The loan was for $280434.00. It was approved on February 17, 2021. The data says it retained 94 jobs. This company name is not the one that customers know it by. The company's most common name is Kitzie's Cafe or Kitzie's Restaurant and Lounge. On the company's website, it appears that the restaurant has two locations, one being Spring Hills and one being in Martinsburg. But the West Virginia business summary states that the Cottle Christi L LLC holdings have owned four Kitzie's and one restaurant named Riverbend Bar and Grill. One location of Kitzie's, Kitzie's Inwood, closed in May 2020 before the restaurant chain received their PPP loan. Another location closed in 2018. 

```{r}

martinsburg_ppp_with_naics %>%
  filter(naics_code == "722511") %>%
  select(amount, name, date_approved, jobs_retained, naics_code) %>%
  arrange(desc(amount))

```
