---
title: "lab_07"
author: "sean mussenden"
date: "8/25/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## About this lab

SOME ADVICE: rvest gives you access to three funtions you will use.

* read_html() = given a url, it readers the html
* html_table() = given some html, it finds all of the <table> objects in it
* html_element() = given some html, you supply one or more tags (like p or ul) OR and xpath expression. An example:
  my_stuff <- url %>%
  read_html() %>%
  html_element('p') # this finds all the <p> tags
* html_text() - contents of any tag

To complete this lab, you need to:
* write code in empty codeblocks provided to answer questions included (look for **Q**).
* write out the answer in the form of a complete sentence in the space given (look for **A**).

When you are finished, commit changes and push to your personal GitHub repo, then submit the URL to this document on ELMS.

## Load libraries and establish settings
**Task**: Load rvest, janitor and the tidyverse
```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse,rvest,janitor
library(tidyverse)
library(rvest)
library(janitor)

```


Q1. How many individual cases has the U.S. Department of Justice filed against people accused of fraud related to the PPP or EIDL loan program, according to the DOJ website listing those cases: https://www.justice.gov/criminal-fraud/cares-act-fraud?  An example of one case that appears on this page is "U.S. v. Richard Ayvazyan et al". To answer this question, you will need to use rvest to scrape the data on this single webpage into a dataframe that lists only case names. Hint: you will need to make use of html_elements() and html_text() -- a function that extracts text inside of an html tag -- for this.
**A1**. There are 101 individual cases filed against people accursed of fraud related to the PPP or EIDL loan program, according to the DOJ website. 

```{r}

cases_url <- "https://www.justice.gov/criminal-fraud/cares-act-fraud"

doj_cases <- cases_url %>%
  read_html() %>%
  html_elements('li b') %>%
  html_text()

cases_dataframe <- as.data.frame(doj_cases)


```

Q2. In how many individual judicial districts has the U.S. Department of Justice filed cases against people accused of fraud related to the PPP or EIDL loan program, according to the DOJ website listing those cases: https://www.justice.gov/criminal-fraud/cares-act-fraud?  Note: an example of a judicial district is "Southern District of Florida". You will need to use rvest scrape the data on this single webpage into a dataframe.
**A2** There are a total of 31 judicial districts where the Department of Justice has filed cases against people accused of fraud related to the PPP or EIDL loan program. 

```{r}

district_url <- "https://www.justice.gov/criminal-fraud/cares-act-fraud"

total_districts <- district_url %>%
read_html() %>%
  html_elements('i') %>%
  html_text()

districts_dataframe <- as.data.frame(total_districts)

```

Q4. The website (https://www.justice.gov/criminal-fraud/cares-act-fraud) shows that the DOJ filed more cases in the Southern District of Florida than in any other district. One of those cases was filed against someone named "Diamond Blue Smith". Who is Smith, and what was he accused of, according to the criminal complaint? If you were an editor, would you have assigned a reporter to write a story about this case when the complaint was filed in court? Why or why not?
**A4** Smith is a rapper. He was arrested for his “alleged participation in a scheme to file fraudulent loan applications seeking more than $24 million in forgivable Paycheck Protection Program (PPP) loans guaranteed by the Small Business Administration (SBA) under the Coronavirus Aid, Relief, and Economic Security (CARES) Act,” according to the Department of Justice complaint. The Miami Herald reported this information in April 2021. I would absolutely assign a reporter to write a story about this case when the complaint was filed in court. This is a relatively public figure in the the Southern Florida community and it is a big deal that he was accursed of more than $24 million in fradulent PPP loans. 

Q5. In what percentage of all judicial districts has the U.S. Department of Justice filed cases cases against people accused of fraud related to the PPP or EIDL loan program, according to the DOJ website listing those cases: https://www.justice.gov/criminal-fraud/cares-act-fraud? In answering this question, you should also produce a list of judicial districts where DOJ has NOT filed a case, according to this site.  Note: to answer this question, you will need to scrape a table of all district courts on this up-to-date Wikipedia page under the heading "Active Courts": https://en.wikipedia.org/wiki/List_of_United_States_district_and_territorial_courts  
**A5** The Department of Justice has filed PPP or EIDL cases in roughly 33 percent of judicial districts. 


```{r}

district_courts_url <- "https://en.wikipedia.org/wiki/List_of_United_States_district_and_territorial_courts"

active_courts  <- district_courts_url %>%
  read_html() %>%
  html_table()

active_courts <- active_courts[[3]] %>%
  select(Region)  

no_doj_cases <- active_courts %>% anti_join(districts_dataframe, by=c("Region"="total_districts"))

answer = (31/94)*100

```
Q6 What might explain why, according to this site, the DOJ has filed PPP-related fraud cases in less than half of the country's judicial districts?
**A6** The DOJ has a very slow enforcement process for determining and prosecuting fraud in different districts. It has to coordinate with agencies across the country. It also did not establish the fraud tax force until the program had already been in place for a year, and the displayed PPP enforcement scheme says that agencies catch fraud best right when an applicant fills out an application. With that said, it is working backwards to determine large amounts of this fraud based on its traditional layout for catching fraudsters. 

Q7. Which state had the most approved PPP loans per 100,000 population? [This web page](https://dwillis.github.io/jour472files/ppp_scraping_example/index.html) has links to 52 individual web pages, one for each state (plus Washington, D.C. and Puerto Rico). Each of those web pages contains a one-row html table that has the name of the state, the number of approved loans, and the 2019 population. Here's an example for [Alabama](https://dwillis.github.io/jour472files/ppp_scraping_example/states/alabama.html). You'll need to loop over the individual state urls, scrape each individual page and combine the information on each page into a single dataframe to answer this question. Don't forget to calculation the loans per 100,000.
**A7**
```{r}

# Create an empty dataframe to hold results

all_states <- "https://dwillis.github.io/jour472files/ppp_scraping_example/index.html"

ppp_loans_per_capita <- tibble()

for(row_number in 1:nrow(all_states)) {

    # Keep only the row for a given row number, get rid of every other row
    each_row_df <- all_states %>%
      slice(row_number)

    # Define url of page to get
    url <- each_row_df$sector_url

    # Define id of table to ingest
    xpath_employment_table <- paste0('//*[@id="',each_row_df$sector_xpath_id,'"]')

    # Get employment table from each page by going to each url defined above, reading in the html with read_html(), extracting the table with the id generated by the xpath code using html_elements), and then turning the html into a proper dataframe using html_table().  The dataframe is in a nested list, which we'll have to extract in the next step.
    employment_info <- url %>%
      read_html() %>%
      html_elements(xpath = xpath_employment_table) %>%
      html_table()

    # Grab the dataframe out of the list (it's the first and only element inside the list); clean up the field names with clean_names(); use slice(2) to keep only the second row; use bind_cols() to append the sector code and name to this table; turn jun_2021 column into a proper number, and rename it.  Then select only three columns we need.
    employment_info <- employment_info[[1]] %>%
      clean_names() %>%
      slice(2) %>%
      bind_cols(each_row_df) %>%
      mutate(nov_2021 = parse_number(nov_2021)) %>%
      rename(nov_2021_employees = nov_2021) %>%
      select(sector,description,nov_2021_employees)

    # Bind each individual employment info table to our employment_by_sector_all dataframe
    employment_by_sector_all <- employment_by_sector_all %>%
      bind_rows(employment_info)

}

# Display the completed dataframe
ppp_loans_per_capita

```
